Interpolants, Error Bounds, and Mathematical Software for Modeling and Predicting Variability in Computer Systems

### 2020-07-29 3-Disseration/1-apps.tex
Update.

>>>>

\section{Organization of this Document}

This thesis is comprised of work from multiple conference and journal
papers. Chapters \ref{ch:apps} and \ref{ch:algs} share the broader
context for this work and derive from many papers
\cite{lux2018predictive,lux2020effective,lux2020interpolation}. Chapter
\ref{ch:naive} appears in \cite{lux2018predictive}. Chapter
\ref{ch:boxes} appears in \cite{lux2018novel}. Chapter \ref{ch:strong}
appears in \cite{lux2018nonparametric}. Chapter \ref{ch:error} is to
appear in \cite{lux2020interpolation}. Finally Chapter
\ref{ch:splines} was presented in part in \cite{lux2020algorithm} and
is to be submitted in full to ACM Transactions on Mathematical
Software (TOMS).

### 2020-07-29 3-Disseration/6-error.tex
Update.

This chapter presents theoretical results bounding the error of
(piecewise) linear interpolation. The error analysis relies on linear
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly apply
to Delaunay interpolation; and (3) multiple other interpolants in this
paper compute predictions as convex combinations of observed function
values, which may allow for straightforward extensions of this error
bound.

@@

Table \ref{table:avg-performance} summarizes results across the four
test data sets with real-valued ranges. The interpolants discussed in
this paper produce the \textit{best} approximations roughly one third
of the time, and produce competitive approximations for almost all
data sets. These test problems are almost certainly
\textit{stochastic} in nature, but the high dimension leads to greater
data sparsity and model construction cost, making interpolation more
competitive.

@@

The various interpolants discussed in this paper have been
demonstrated to effectively approximate multivariate phenomena up to
$30$ dimensions. The underlying constructions are theoretically
straightforward, interpretable, and yield reasonably accurate
predictions. Most of the interpolants' computational complexities make
them particularly suitable for applications in even higher
dimension. The major benefits of interpolation are seen when only a
small number of approximations ($\leq 1000$) are made from data and
when there are relatively few data points for the dimension (for
empirical results presented, $\log_d n \leq 5$). These findings
encourage broader application of interpolants to multivariate
approximation in science.

>>>>

This chapter presents theoretical results bounding the error of
(piecewise) linear interpolation. The error analysis relies on linear
interpolation for three reasons: (1) second order results can be
obtained utilizing a Lipschitz constant on the gradient of a function,
rather than standard Lipschitz bounds; (2) the results directly apply
to Delaunay interpolation; and (3) multiple other interpolants in this
work compute predictions as convex combinations of observed function
values, which may allow for straightforward extensions of this error
bound.

@@

Table \ref{table:avg-performance} summarizes results across the four
test data sets with real-valued ranges. The interpolants discussed in
this work produce the \textit{best} approximations roughly one third
of the time, and produce competitive approximations for almost all
data sets. These test problems are almost certainly
\textit{stochastic} in nature, but the high dimension leads to greater
data sparsity and model construction cost, making interpolation more
competitive.

@@

The various interpolants discussed in this work have been
demonstrated to effectively approximate multivariate phenomena up to
$30$ dimensions. The underlying constructions are theoretically
straightforward, interpretable, and yield reasonably accurate
predictions. Most of the interpolants' computational complexities make
them particularly suitable for applications in even higher
dimension. The major benefits of interpolation are seen when only a
small number of approximations ($\leq 1000$) are made from data and
when there are relatively few data points for the dimension (for
empirical results presented, $\log_d n \leq 5$). These findings
encourage broader application of interpolants to multivariate
approximation in science.
